Optimizing Abstract Abstract Machines

### 2012-11-13 paper/paper.tex
Rewording in literal data abstraction.

The $\machstep$ relation is most naturally written as a function from
a state to a set of all next states. We want to find a fixpoint of
$\machstep$ on the state space, however, so we must lift it to work on
sets of states. The lifting is a natural one:

@@

The number of syntactic uses of {\tt cons} versus implicit uses via
literal lists is dramatically smaller in typical Scheme programs. We
use the above abstraction for these syntactic uses, but choose to
interpret literal lists as not always sugar for cascading {\tt
  cons}es. In particular, if a list literal is ``too big'' (in our
case length $>$ 1), we interpret the list as a circular data structure;
the right address points back to the {\tt cons} itself, and the left
address points to all of the elements of the list. We take this
farther and join the list elements together in with a coarse
type-based abstraction. Quotation is special because it cannot
introduce function values, which is important to enhancing our
technique's soundness to conceptual complexity ratio. If we join two
values of different type together, we don't get ``anything,'' which
has complex meaning in higher-order languages (Shivers' escape
semantics) and is overly approximate. We instead get ``any quotable
value,'' which has much simpler semantics.

@@

Besides the original AAM work, the analysis most similar to that
presented in section~\ref{sec:aam} is the infinitary control-flow
analysis of Nielson and Nielson~\cite{dvanhorn:nielson-nielson-popl97}
and the unified treatment of flow analysis by Jagannathan and
Weeks~\cite{dvanhorn:jagannathan-weeks-popl95}.  Both are
parameterized in such a way that in the limit, the analysis is
equivalent to an interreter for the language, just as is the case
here.  What is different is that both give a constraint-based
formulation of the abstract semantics rather than a finite machine
model.

>>>>

The $\machstep$ relation is most naturally written as a function from
a state to a set of all next states. To calculate $eval$, we find a fixed point of
$\machstep$ lifted to step a set of states. The lifting is a natural one:

@@

The number of syntactic uses of {\tt cons} versus implicit uses via
literal lists is dramatically smaller in typical Scheme programs. We
use the above abstraction for these syntactic uses, but choose to
interpret literal lists as not always sugar for cascading {\tt
  cons}es. In particular, if a list literal is ``too big'' (in our
case length $>$ 1), we interpret the list as a circular data
structure; the right address points back to the {\tt cons} itself, and
the left address points to all of the elements of the list. We take
this farther and join the list elements together in with a coarse
type-based abstraction. In effect, large lists or vectors of literal
numbers become unbounded lists/vectors of ``number.'' Heterogeneous
data is combined to just ``data,'' rather than a union of ``number'',
``string'', etc. since primitives are interpreted on every combination
of values that flow to them, and unions lead to more primitive
interpretations.

Quotation is special because it cannot introduce function values,
which is important to enhancing our technique's soundness to
conceptual complexity ratio. If we join two values of different type
together, we don't get ``anything,'' which has complex meaning in
higher-order languages (Shivers' escape semantics) and is overly
approximate. We instead get ``any quotable value,'' which has much
simpler semantics.

@@

Besides the original AAM work, the analysis most similar to that
presented in section~\ref{sec:aam} is the infinitary control-flow
analysis of Nielson and Nielson~\cite{dvanhorn:nielson-nielson-popl97}
and the unified treatment of flow analysis by Jagannathan and
Weeks~\cite{dvanhorn:jagannathan-weeks-popl95}.  Both are
parameterized in such a way that in the limit, the analysis is
equivalent to an interpreter for the language, just as is the case
here.  What is different is that both give a constraint-based
formulation of the abstract semantics rather than a finite machine
model.
